---
title: "HIVCAre"
author: "Clinical Data Science Group der DKTK CCP"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    css: style.css
subtitle: "Epidemiology and Inpatient Care Characteristics of HIV-positive Cancer
  Patients: Exploration of Real World Data from Four German University Hospitals"
editor_options:
  chunk_output_type: inline
---

```{r setup}

knitr::opts_chunk$set(
  fig.path = "images", # use only for single Rmd files
  collapse = TRUE,
)


# (Re-)Run analysis if needed
#source("./src/MAIN.R")

# Get workspace objects from analysis scripts
#load("./output/outputdata.RData")

# Check and load packages needed for report
pacman::p_load(DiagrammeR,
               dplyr,
               forcats,
               ggplot2,
               ggsankey,
               ggstatsplot,
               ggthemr,
               gt,
               lubridate,
               purrr,
               stringr,
               tidyr,
               readr)

```


### Objectives of descriptive analysis and data exploration

* A)  Comparison of all subgroups
      + A1) Sample sizes and Demography (Sex, Age)
      + A2) Case count per patient
      + A3) Mean length of stay per patient
      + A4) Spatial distribution (Counts of patients from different postal code areas)

* B)  HIVandCancer
      + B1) Primary cancer diagnosis subgroups and occurrence per year
      + B2) Age at admission separated for diagnosis subgroups and in 4-year time periods

* C)  HIVandCancer vs. CancerOnly
      + C1) Age at presumed cancer onset
      + C2) Metastasis occurrence
      + C3) Time-to-event: Cancer onset to metastasis onset
      + C4) Count of distinct cancer entities
      + C5) Therapy modalities
      + C6) Count of chemotherapy sessions
      + C6) Discharge reasons (Death, Home)
      + C7) Occurrence of complications (ventilation, dialysis) after chemotherapy
      + C8) AIDS after chemotherapy
      + C9) Death after chemotherapy

* D)  HIVandCancer vs. HIVOnly
      + D1) Age at presumed HIV onset
      + D2) AIDS occurrence
      + D3) Time-to-event: HIV onset to AIDS onset
      
* E)  Sankey diagram: Sequence in presumed HIV / Cancer / Metastasis / AIDS onset

* F)  Sankey diagram: Sequence of cancer therapy modalities


### Objectives of inference analysis

** No explicit inference analysis goals **





### Analysis data: Sources, Requirements, Assumptions

#### Analysis data sources

* Datenintegrationszentrum (DIZ)


#### Analysis data description

```{r table_RequiredData}

df_Report_RequiredData <- tibble(Description = character(),
                                 DataFormat = character(),
                                 Source = character(),
                                 DataElements = character()) %>%
                                 #----------------------------------------------
                                 add_row(Description = "Case data",
                                         DataFormat = ".csv-file",
                                         Source = "DIZ",
                                         DataElements = "CasePseudonym<br>
                                                         PostalCodeShort<br>
                                                         YearOfBirth<br>
                                                         Sex<br>
                                                         AdmissionDate<br>
                                                         DischargeDate<br>
                                                         DischargeReason<br>
                                                         AdmissionAge") %>%
                                 add_row(Description = "ICD codes per case",
                                         DataFormat = ".csv-file",
                                         Source = "DIZ",
                                         DataElements = "CasePseudonym<br>
                                                         DiagnosisType<br>
                                                         ICDVersion<br>
                                                         ICDCode<br>
                                                         Localization<br>
                                                         SecondaryICDCode<br>
                                                         SecondaryLocalization") %>%
                                 add_row(Description = "OPS codes per case",
                                         DataFormat = ".csv-file",
                                         Source = "DIZ",
                                         DataElements = "CasePseudonym<br>
                                                         OPSVersion<br>
                                                         OPSCode<br>
                                                         Localization<br>
                                                         OPSDate") %>%
                                 add_row(Description = "Case to patient mapping",
                                         DataFormat = ".csv-file",
                                         Source = "DIZ",
                                         DataElements = "CasePseudonym<br>
                                                         PatientPseudonym")

(tbl_Report_RequiredData <- df_Report_RequiredData %>%
                                gt(rowname_col = "Data Description") %>%
                                tab_caption("Required analysis data") %>%
                                tab_stubhead(label = "Data Description") %>%
                                cols_label(DataFormat = "Data Format",
                                           Source = "Source",
                                           DataElements = "Data Elements")) %>%
                                fmt_markdown(columns = everything())            # Enable markdown formatting in all columns

```


### Meta data description







### Assumptions and definitions in analysis rational


#### General ICD coding

It is assumed that ICD Coding was performed abiding by established coding manuals (e.g. ZI HIV coding manual)

#### ICD cancer coding

The following ICD codes were used to identify patients presumed to have a cancerous disease in their history:

```{r table_Definitions, echo = FALSE}

df_Report_CancerCodes <- tibble(ICDCode = character(),
                                Interpretation = character()) %>%
                                #-----------------------------------------------
                                add_row(ICDCode = "Starting with C", Interpretation = "All malignant neoplasms (primary and secondary)") %>%
                                add_row(ICDCode = "Starting with D0", Interpretation = "In-situ neoplasms") %>%
                                add_row(ICDCode = "D37 - D48", Interpretation = "Neoplasms of uncertain / unknown behaviour") %>%
                                add_row(ICDCode = "Z08", Interpretation = "Follow-up examination after treatment for malignant neoplasms") %>%
                                add_row(ICDCode = "Z85, Z92.6", Interpretation = "Personal history of cancer / cancer therapy")

(tbl_Report_CancerCodes <- df_Report_CancerCodes %>%
                              gt(rowname_col = "ICDCode") %>%
                              tab_caption("ICD-10 cancer codes used in analysis") %>%
                              tab_stubhead(label = "ICD-10 codes") %>%
                              cols_label(Interpretation = "Interpretation"))
                              
```

ICD codes for metastases (C77, C78, C79) are **not** considered cancer codes and are treated separately.


#### Notes on inclusion of presumed cancer patients

In order to get a sufficient degree of integrity within the explored data, we include only patients for whom a cancer code with diagnosis type *Hauptdiagnose* is present.
We consider the first occurrence of such a code within a patient's documented history as the main cancer entity, that the respective patient is treated for.
Furthermore, we refer to the admission date belonging to this code as the date of cancer onset.

#### Notes on presumption of AIDS manifestation

* HIV code before or at the same time as AIDS defining code
* Certain AIDS-defining cancer codes may occur ahead in time of HIV code occurrence. Patients in question would not be presumed to have AIDS.


### Attrition Plot

```{r plot_Attrition, echo = FALSE}

plot_Output_Attrition

```



### Notes on concerns, limitations and potential bias sources


```{r table_Concerns, echo = FALSE}

df_Report_Concerns <- tibble(Concern = character(),
                             PotentialImpact = character(),
                             Comment = character()) %>%
                             #--------------------------------------------------
                             add_row(Concern = "Disease onsets may be assumed wrongfully (as too late). Therefore, calculated onset intervals may be incorrect.",
                                     PotentialImpact = "Major",
                                     Comment = "") %>%
                             add_row(Concern = "If a patient has more than one distinct cancer entity, therapy measures can not be linked to different entities.",
                                     PotentialImpact = "Major",
                                     Comment = "") %>%
                             add_row(Concern = "Age at Admission not exact due to lack of exact birth date.",
                                     PotentialImpact = "Minor",
                                     Comment = "") %>%
                             add_row(Concern = "Occurrence of implausible HIV coding.",
                                     PotentialImpact = "Minor",
                                     Comment = "") %>%
                             add_row(Concern = "Postal code of patient taken from first admission, although moving between admissions is imaginable.",
                                     PotentialImpact = "Minor",
                                     Comment = "")
  

(tbl_Report_Concerns <- df_Report_Concerns %>%
                              gt(rowname_col = "Concern") %>%
                              tab_caption("Concerns") %>%
                              tab_stubhead(label = "Concern") %>%
                              cols_label(PotentialImpact = "Estimated potential impact",
                                         Comment = "Comment"))
```




### Stratification

```{r plot_Stratification, echo = FALSE}

plot_Output_Stratification

```





```{r table_Demography, echo = FALSE}

df_Output_Demography %>%
    gt()

```

